Free Agency in Baseball: MLB Teams Aren’t Getting The Player They Bargained For

### 2022-11-28 manuscript/manuscript.tex
Update manuscript.tex

These analyses fall victim to the lofty standards of statistical significance. However, it can still be seen that the same 
large drops can be seen in these value-added statistics. Both categories fell over 40% from 2021 to 2022, even if neither was 
significant at the level alpha = 0.10. It should be noted, however, that both P-values came extremely close to being significant 
at that level. While our ANOVA may not have produced any definitive conclusions here, we can still see that our players, on 
average, added nearly six more runs to their teams in 2021 than in 2022.

@@

Arguably the most troubling finding for our players, even more so than their dismal health, was their inability to meet 
expectations when they were on the field. Our fourth and final table contained statistical categories whose drop-offs could 
not be explained by not playing as many games in 2022. Every category decreased in some way except for strikeout percentage, 
with slugging percentage, home run percentage, and walk percentage decreasing significantly. It seems as though between power, 
contact, speed, and discipline, power is the performance tool that appears to have evaporated the most in our players. The drop 
in slugging percentage and home run percentage matches up with the drop in other power-related categories from table one. 
Hitting power is a complicated concept. Gratuitous strength doesn’t always correlate with good power, but it certainly does 
play a part. Perfect swing form and unimpeachable timing ability are inarguably more important when it comes to launching the 
baseball- strength means nothing if you cannot get the barrel of the bat on the ball. A baseball swing can be thought of as a 
symphony of moving body parts. If a batter develops even the slightest hitch in their swing, the entire orchestra is thrown out 
of balance and the player stands no chance of hitting the ball squarely. These hitches usually develop from, coincidentally, 
nagging injuries to the back, neck, and extremities. With so many of the players in our sample battling long-term ailments in 
2022, it is not difficult to see how their power was essentially zapped when compared to the season before. These weeks-long 
bouts with injury can cause major performance slumps, draining a player’s confidence and setting them up for failure. Baseball 
is partially a mental game- step up to the plate with a pessimistic mindset, and you’re almost guaranteed to fail. Early-season 
struggles, especially with a brand-new team, can easily spiral into season-long, emotionally torturous campaigns that leave 
fans wondering what exactly the team signed up for when they shelled out millions of dollars.

>>>>

These analyses fall victim to the lofty standards of statistical significance. However, the same large drops can be seen in these 
value-added statistics. Both categories fell over 40% from 2021 to 2022, even if neither was significant at the level alpha = 0.10. 
It should be noted, however, that both P-values came extremely close to being significant at that level. While our ANOVA may not 
have produced any definitive conclusions here, we can still see that our players, on average, added nearly six more runs to their 
teams in 2021 than in 2022.

@@

Arguably the most troubling finding for our players, even more so than their dismal health, was their inability to meet 
expectations when they were on the field. Our fourth and final table contained statistical categories whose drop-offs could 
not be explained by not playing as many games in 2022. Every category decreased in some way except for strikeout percentage, 
with slugging percentage, home run percentage, and walk percentage decreasing significantly. It seems as though between power, 
contact, speed, and discipline, power is the performance tool that appears to have evaporated the most in our players. The drop 
in slugging percentage and home run percentage matches up with the drop in other power-related categories from table one. 
Hitting power is a complicated concept. Gratuitous strength doesn’t always correlate with good power, but it certainly does 
play a part. Perfect swing form and unimpeachable timing ability are inarguably more important when it comes to launching the 
baseball- strength means nothing if you cannot get the barrel of the bat on the ball. A baseball swing can be thought of as a 
symphony of moving body parts. If a batter develops even the slightest hitch in their swing, the entire orchestra is thrown out 
of balance and the player stands no chance of hitting the ball squarely. These hitches usually develop from, coincidentally, 
nagging injuries to the back, neck, and extremities. With so many of the players in our sample battling long-term ailments in 
2022, it is not difficult to see how their power was vastly decreased when compared to the season before. These weeks-long 
bouts with injury can cause major performance slumps, draining a player’s confidence and setting them up for failure. Baseball 
is partially a mental game- step up to the plate with a pessimistic mindset, and you’re almost guaranteed to fail. Early-season 
struggles, especially with a brand-new team, can easily spiral into season-long, emotionally torturous campaigns that leave 
fans wondering what exactly the team signed up for when they shelled out millions of dollars.
